From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study

### 2015-08-21 WOLFHPC15/article/conclusion.tex
minor diff

This paper has presented the domain specific language MSL and its compiler MSCAC designed to produce a parallel (data and hybrid) component-based runtime of an overall multi-stencil program, \ie a mesh-based numerical simulation reduced to explicit schemes. After details on the language and its compiler, MSL has been evaluated by the description of a real case numerical simulation: the shallow water equations. The component-based data parallelization of the simulation has been compared to a pure SkelGIS parallelization, and has shown improved execution times as well as a promising scalability. Those results demonstrate that component-based runtimes may be relevant back-end codes for DSLs as they do not introduce performance damage. Moreover, components bring software engineering benefits such as separation of concern, code re-use, improving maintainability.

Although MSL is a promising case study from DSLs to component-based runtimes, many works in progress aims at improving this first contribution. First, to more clearly show the improvement of DSLs maintainability using component-based back-end, an alternative \texttt{DDS} component is under study, using Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}. In addition to this, alternatives for the \texttt{Computations} component, computed by MSC, are under study such as a dump to an OpenMP~\cite{660313} code or the use of dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}. This last work on dynamic schedulers could also improve the expressivity of task parallelism in MSL, as it has been discussed in Section~\ref{sect:eval}, and thus it could bring interesting performance for the hybrid parallelization.

>>>>

This paper has presented the domain specific language MSL and its compiler MSCAC designed to produce a parallel (data and hybrid) component-based runtime of an overall multi-stencil program, \ie a mesh-based numerical simulation reduced to explicit schemes. After details on the language and its compiler, MSL has been evaluated by the description of a real case numerical simulation: the shallow water equations. The component-based data parallelization of the simulation has been compared to a pure SkelGIS parallelization, and has shown improved execution times as well as a promising scalability. Those results demonstrate that component-based runtimes may be relevant back-end codes for DSLs as they do not introduce performance damage. Moreover, components bring software engineering benefits such as separation of concerns, code re-use, improving maintainability.

Although MSL is a promising case study from DSLs to component-based runtimes, many works in progress aims at improving this first contribution. First, to more clearly show the improvement of DSLs maintainability using component-based back-end, an alternative \texttt{DDS} component is under study, using Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}. In addition to this, alternatives for the \texttt{Computations} component, computed by MSC, are under study such as a dump to an OpenMP~\cite{660313} code or the use of dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}. This last work on dynamic schedulers could also improve the expressivity of task parallelism in MSL, as discussed in Section~\ref{sect:eval}. Thus, it could bring interesting performance for the hybrid parallelization.

### 2015-08-21 WOLFHPC15/article/mscac.tex
minor diff

As illustrated in Figure~\ref{fig:mscac}, the compiler, named MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of three different phases. The first phase \emph{parses} the MSL description into an intermediate representation (IR). The second phase called MSC transforms the computation part ($\Gamma$) of the IR in order to automatically introduce parallelism in the overall numerical simulation.
Finally, the third phase dumps the IR to an actual implementation of the parallel pattern of the simulation.

>>>>

As illustrated in Figure~\ref{fig:mscac}, the compiler, named MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of three different phases. The first phase \emph{parses} the MSL description into an intermediate representation (IR). The second phase called MSC transforms the computation part ($\Gamma$) of the IR into a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393} to introduce parallelism in the overall numerical simulation.
Finally, the third phase dumps the IR to an actual implementation of the parallel pattern of the simulation.
